---
---
---
title: "Photosynthesis lab"
author: "Senanu Spring-Pearson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=TRUE, echo=FALSE}
knitr::opts_knit$set(global.device = TRUE, echo=TRUE) # leave this alone
```

# Getting Started

As you have done before, please make sure you can successfully knit this document before you make any changes to it.

Take out my name in the headers (line 3) and put your name in it's place.

# Photosynthesis lab

Make a graph showing the absorption spectrum of the pigments in elodea. You should:

1. Save a google sheets data table with the headers "wavelength" and "absorbance". Remember not to include extraneous characters like parentheses or spaces.
2. Read the data into this document and save it in a variable called `dat`.
3. Use the `plot()` command to generate a figure showing how the absorbance varies with the wavelength.
  1. remember to tell R which data column you want on which axis. You can do this with the `x=dat$variable1` and `y=dat$variable2` arguments to the `plot()` command.
  2. Add axis labels with `xlab="my X-axis label"` and similarly for `ylab=`.
  3. Connect the points with `type='l'` (that's the lower-case letter 'l' which stands for 'lines'). As with the `xlab` and `ylab` arguments, this should be an argument to the `plot()` command.
4. When the code chunk below works, change "eval=FALSE" to "eval=TRUE"
5. Knit this document and turn it in.

```{r absorbancePlot, eval=FALSE, echo=TRUE, include=TRUE}
dat <-
plot(  )
```
